---
title: "UK interest rate assumption"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)


```


This note presents the possible options for the interest rate assumptions for the project.


We need long-term interest rate data for a long period of time so that we can calculate annuity values. In addition we need the term structure of interest rates to be able to illustra non-parallel changes in interst rates.

# Long term interest rates

For the long term interest rates we can follow Willis Tower UK long-term statistics. Watson^[https://www.willistowerswatson.com/-/media/WTW/Insights/2019/08/uk-long-term-statistics-2019.pdf]. According to this we can use the yields of UK Consols^[https://en.wikipedia.org/wiki/Consol_(bond)] which are available from 1753 till 2015. From 2015 onwards we can use the yield of 20 year UK gilt. These data are available from the Bank of England^[https://www.bankofengland.co.uk/statistics/research-datasets] in the so-called "A millennium of macroeconomic data".


```{r readData}
lngRate <- read_csv("UKlongtermrate.csv")
lngRate <- lngRate %>%
  mutate(date = dmy(paste("1", lngRate$Month, lngRate$Year, sep = "-")))
```

This data is ploted below

```{r plotLongTerm}
ggplot(lngRate) + geom_line(aes(x = date, y = rate)) + ylim(0,NA) + 
  geom_line(aes(x = date, y = smooth.spline(lngRate$rate, spar = 0.4)$y), colour = "blue")

```


# Term structure

The data on the term structure of the interes rate can be obtained from https://www.bankofengland.co.uk/statistics/yield-curves. and is available from 1985







